### 2025-04-24 docs/tex_src/main.tex
notes

\documentclass[a4paper,10pt]{article}
\usepackage[utf8]{inputenc}
\newcommand{\myparagraph}[1]{\paragraph{#1}\mbox{}\\}
\usepackage[a4paper,margin=1.5cm]{geometry} %Sets the page geometry
\usepackage{url}
\usepackage{dirtytalk}
\usepackage{graphicx} % Package for \includegraphics
\usepackage{wrapfig} % Figure wrapping
\usepackage[T1]{fontenc} % Output font encoding for international characters
\setlength{\parskip}{1em} % Set space when paragraphs are used
\usepackage{amssymb}
\usepackage{amsmath}
\usepackage{tcolorbox}
\usepackage{mathtools}
\usepackage{enumerate}
\usepackage{algorithm}
\usepackage{algorithmicx}
\usepackage{algpseudocode}

% Lets you use \blankpage to make a blank page
\newcommand{\blankpage}{
	\newpage
	\thispagestyle{empty}
	\mbox{}
	\newpage
}

% Self Explanatory
\newtheorem{theorem}{Theorem}[section]
\newtheorem{definition}{Definition}[section]
\newtheorem{corollary}{Corollary}[theorem]
\newtheorem{lemma}[theorem]{Lemma}
\newtheorem{example}{Example}[section]
\newtheorem{proof}{Proof}[section]

% Other
\DeclarePairedDelimiter\floor{\lfloor}{\rfloor} %Floor function

\begin{document}
\begin{flushleft}
	\section{P2P Network Design}
	\begin{itemize}
		\item $N$ - nodes in network
		\item $L$ - layers in model
		\item $l_i = (x,y) : x,y \in [0,L); y > x$ - range of layers node $i$ has
		\item $\ell _{ij}$ - latency from node $i$ to node $j$
		\item $c_i$ - computational cost of node $i$
		\item $p_i$ - preload cost of node $i$
		\item $e_i$ - embedding cost of node $i$
	\end{itemize}

	\section{Optimal Path Finding}
	The P2P network is constructed as a directed graph. Edges are compatiable nodes with the next layers of the model.
	We know that this is a DAG becasue the layers must be preformed sequentially. The weights of the edges are the sum of latency and computational cost.
	The optimal path can be found using a topological sort. The start node is the users computer, and the terminal nodes are all nodes that have layer N.
	The optimal path connects the start node to a terminal node.

	\subsection{Topological Sort}
	\begin{algorithm}
		\caption{Topological Sort of a Directed Acyclic Graph (DAG)}
		\begin{algorithmic}[1]
			\Require{A directed acyclic graph $G = (V, E)$}
			\Ensure{A linear ordering of vertices such that for every directed edge $(u, v)$, vertex $u$ comes before vertex $v$}

			\State $\text{result}[]$ \Comment An empty list to store the sorted vertices
			\State $\text{visited}[\text{false}; V]$ \Comment A map to track visited vertices
			\State $\text{temp}[\text{false}; V]$ \Comment A map to track vertices in the current recursion stack

			\For{each vertex $v \in V$}
			\If{$\text{visited}[v] = \text{false}$}
			\State DFS-Visit($G$, $v$, $\text{visited}$, $\text{temp}$, $\text{result}$)
			\EndIf
			\EndFor

			\State \Return Reverse($\text{result}$)

			\Function{DFS-Visit}{$G$, $u$, $\text{visited}$, $\text{temp}$, $\text{result}$}
			\State $\text{temp}[u] \gets \text{true}$ \Comment{Mark current vertex as being processed}

			\For{each vertex $v$ such that $(u, v) \in E$}
			\State DFS-Visit($G$, $v$, $\text{visited}$, $\text{temp}$, $\text{result}$)
			\EndFor

			\State $\text{temp}[u] \gets \text{false}$ \Comment{Mark $u$ as processed}
			\State $\text{visited}[u] \gets \text{true}$ \Comment{Mark $u$ as visited}
			\State Append $u$ to $\text{result}$
			\EndFunction
		\end{algorithmic}
	\end{algorithm}
	\subsection{Dynamic Programming}
\end{flushleft}
\end{document}