### 2025-05-01 docs/paper/main.tex
paper outline

\documentclass[preprint,twoside,11pt]{article}

\usepackage{blindtext}

% Any additional packages needed should be included after jmlr2e.
% Note that jmlr2e.sty includes epsfig, amssymb, natbib and graphicx,
% and defines many common macros, such as 'proof' and 'example'.
%
% It also sets the bibliographystyle to plainnat; for more information on
% natbib citation styles, see the natbib documentation, a copy of which
% is archived at http://www.jmlr.org/format/natbib.pdf

% Available options for package jmlr2e are:
%
%   - abbrvbib : use abbrvnat for the bibliography style
%   - nohyperref : do not load the hyperref package
%   - preprint : remove JMLR specific information from the template,
%         useful for example for posting to preprint servers.
%
% Example of using the package with custom options:
%
% \usepackage[abbrvbib, preprint]{jmlr2e}

\usepackage{jmlr2e}

% Definitions of handy macros can go here

\newcommand{\dataset}{{\cal D}}
\newcommand{\fracpartial}[2]{\frac{\partial #1}{\partial  #2}}

% Heading arguments are {volume}{year}{pages}{date submitted}{date published}{paper id}{author-full-names}

\usepackage{lastpage}
\jmlrheading{23}{2022}{1}{1/21}{9/22}{21-0000}{Author One and Author Two}

% Short headings should be running head and authors last names

\begin{document}

\title{P2P-LLM: Collaborative LLM Inference}

\author{\name Andrew Boessen \email boessena@bc.edu \\
}

\maketitle

\begin{abstract}%   <- trailing '%' for backward compatibility of .sty file
	\blindtext
\end{abstract}

\section{Introduction}

\section{Realted Work}

\subsection{Crowd Computing}

\subsection{P2P Networks}

\section{Peer-to-Peer Network}

\subsection{Path Finding}

\subsection{Revenue Optimization}

\subsection{Network Equilibrium}

\subsection{Strategy-Proofness}

\section{Blockchain}

\subsection{Contracts}

\section{Results}

Here is a citation \cite{chow:68}.

% Manual newpage inserted to improve layout of sample file - not
% needed in general before appendices/bibliography.

\newpage

\appendix
\section{}

\vskip 0.2in
\bibliography{main}

\end{document}